### 2017-11-06 1/exams/m2.tex
Progress on m2.

\begin{center}
  {\Large {\bf CMSC 131A, Midterm 2 (Practice)}\\
    \censor{{\bf SOLUTION}} \ \\
    \ \\
    Fall 2017 \ \\
  }
\end{center}

@@

\vskip 2.5in

\begin{SaveVerbatim}{VerbEnv}  
SOLUTION:
\end{SaveVerbatim}

@@

\begin{verbatim}
;; f :

(define (f x)
  (+ (sqr x) 2))

;; g :

(define (g x)
  (map add1 x))

;; h :

(define (h x y)
  (cond [(empty? x) y]
        [(cons? x)
         (cons (sqr (first x))
               (h (rest x) y))]))

;; i :

(define (i x)
  (sqr (x 5)))

;; j :

(define (j x)
  (lambda (y)
    (+ x y)))

;; k :

(define (k x y)
  (cond [(= y 0) (list (x 0))]
        [else
          (cons (x y) (k x (sub1 y)))]))
\end{verbatim}

@@

\noindent
{\bf Problem 3 (15 points).}
%
Design a program called {\tt avg} that computes the average of a list of numbers.

\newpage 
\noindent
{\bf Problem 4 (15 points).}
%
Here is a parametric data definition for a non-empty list of items:
\begin{verbatim}
;; A [NEListof X] is one of:
;; - (cons X '())
;; - (cons X [NEListof X])
;; Interp: an arbitrarily long but non-empty sequence of items.
\end{verbatim}
Here are two similar programs that operate over non-empty lists:
\begin{verbatim}
;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;

;; smallest : [NEListof Number] -> Number
;; Produce the smallest number in the list
(check-expect (smallest (list 6 3 2 4)) 2)
(define (smallest xs)
  (cond [(empty? (rest xs)) (first xs)]
        [else
          (min (first xs)
               (smallest (rest xs)))]))

;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;

;; longest : [NEListof String] -> String
;; Produce the first longest string in the list
(check-expect (longest (list "hi" "there")) "there")
(check-expect (longest (list "same" "size")) "same")
(define (longest xs)
  (cond [(empty? (rest xs)) (first xs)]
        [else
          (longer (first xs)
                  (longest (rest xs)))]))

;; longer : String String -> String
;; Select the longer of the two strings (s1 if tied)
(check-expect (longer "hi" "there") "there")
(check-expect (longer "same" "size") "same")
(define (longer s1 s2)
  (cond [(< (string-length s1) (string-length s2)) s2]
        [else s1]))
\end{verbatim}

Design an abstraction of the {\tt smallest} and {\tt longest}
functions and redefine {\tt smallest} and {\tt longest} in terms of
it.  Be sure to give the most general parametric signature for your
abstraction function.

\noindent
[Space for problem 4.]

\newpage
\noindent
{\bf Problem 5 (15 points).}  
%
Design a program that takes a list of numbers and produces the count
of positive numbers in the list.  For example, if the list contains 3,
4, -1, 0, -2, and 5, the count of positive numbers is 3.  You may
assume the {\tt{[Listof Number]}} data definition is defined.
%
For full credit, use list abstraction functions.  For partial credit,
follow the template for {\tt{[Listof Number]}}.

@@

@@

@@

@@

\noindent
{\bf Problem 7 (15 points).}
%
A sports tournament in which teams pair off against eachother and
the winner advances to the next level in the tournament can be modelled
with the following data definition:

\begin{verbatim}
;; A Tournament is one of:
;; - String
;; - (make-bracket Tournament Tournament)
;; Interp: a sports tournament where a String is a team name
;; that is still in the tournament and a bracket represents
;; a match between the winner of two tournaments. 
\end{verbatim}

Design a function {\tt most-games} that computes the maximum number of
games any team in the tournament would have to play in order to win
the tournament.

>>>>

\begin{center}
  {\Large {\bf CMSC 131A, Midterm 2}\\
    \censor{{\bf SOLUTION}} \ \\
    \ \\
    Fall 2017 \ \\
  }
\end{center}

@@

(define (f b s) s)
(define (g fs) (fs #false))
(define (h bs ns) 2)
(define (i f) (f 2))
(define (j f xs) (map f xs))
(define (k f g x) (g (f x)))
\end{SaveVerbatim}

@@

\begin{alltt}
;; method : \censor{Number -> Number}

(define (method x)
  (+ (sqr x) 2))

;; raekwon : \censor{[Listof Number] -> Boolean}

(define (raekwon x)
  (ormap positive? x))

;; rza : \censor{[Listof String] [Listof Number] -> [Listof Number]}

(define (rza x y)
  (cond [(empty? x) y]
        [(cons? x)
         (cons (string-length (first x))
               (rza (rest x) y))]))

;; gza : \censor{(String Number -> Number) -> Number}

(define (gza x)
  (foldr x 0 (list "a" "b" "c")))

;; odb : \censor{(Number -> Number) -> (Number -> Number)}

(define (odb x)
  (lambda (y)   
    (/ (- (x (+ y 0.001))
          (x (- y 0.001)))
       (* 2 0.001))))

;; ghostface : \censor{[X Y] [Listof X] [X -> Y] -> [Listof Y]}

(define (ghostface x y)
  (cond [(empty? x) '()]
        [(cons? x)
         (cons (y (first x)) 
               (ghostface (rest x) y))]))
\end{alltt}

@@

\noindent
{\bf Problem 3 (15 points).}
%
Design a program called {\tt w-avg} that computes the weighted average
of a list of numbers and a list of weights.

For example, let's say a class grade is based on two midterms and a
project where the project is worth twice as much as the midterms.  A
student who gets an 80 and a 70 on the midterms and a 90 on the
project would have a weighted average of 82.5: $((1 \times 80) + (1
\times 70) + (2 \times 90)) / 4$, which can be computed with {\tt
  (w-avg (list 80 70 90) (list 1 1 2))}.  You should assume the two
lists have the same length.

@@

;; avg : [Listof Number] -> Number
;; Compute the average of a list of numbers
;; Assume: the list is non-empty
(check-expect (avg (list 3 4 5)) 4)
(check-expect (avg (list 10 20)) 15)
(define (avg lon)
  (/ (sum lon) 
     (length lon)))

;; sum : [Listof Number] -> Number
;; Sum the list of numbers
(check-expect (sum '()) 0)
(check-expect (sum (list 3 4 5)) 12)
(define (sum lon)
  (foldr + 0 lon))

;; Alt: follow [Listof Number] template
\end{SaveVerbatim}

\newpage 
\noindent
{\bf Problem 4 (20 points).}
%
Here is a parametric data definition for a tree of elements:
\begin{verbatim}
;; A [Tree X] is one of:
;; - (make-leaf)
;; - (make-node X [Tree X] [Tree X])
;; Interp: a binary tree that is either empty (a leaf), or non-empty (a node)
;; with an element and two sub-trees.
(define-struct leaf ())
(define-struct node (elem left right))
\end{verbatim}
{\bf Problem 4(a) (10 points).} Design a function, {\tt tree-sum},
that operates on a binary tree of numbers, {\tt [Tree Number]}, and
produces the sum of all the numbers in the tree.

;; tree-sum : [Tree Number] -> Number
;; Sum all the numbers in the tree
(check-expect (tree-sum (make-leaf)) 0)
(check-expect (tree-sum (make-node 7 (make-node 1 (make-leaf) (make-leaf)) 
                                     (make-leaf)))
              8)
(define (tree-sum bt)
  (cond [(leaf? bt) 0]
        [(node? bt)
         (+ (node-elem bt)
            (tree-sum (node-left bt))
            (tree-sum (node-right bt)))]))

@@

\noindent
{\bf Problem 4(b) (10 points).}
Here is a function that produces a list of all the elements in a tree:
\begin{verbatim}
;; tree-elems : [X] . [Tree X] -> [Listof X]
;; Produce a list of all elements in tree going top-down, left-to-right
(check-expect (tree-elems (make-leaf)) '())
(check-expect (tree-elems (make-node 7 
                                     (make-node 9 (make-leaf) (make-leaf))
                                     (make-node 2 (make-leaf) (make-leaf))))
              (list 7 9 2))
(define (tree-elems bt)
  (cond [(leaf? bt) '()]
        [(node? bt)
         (cons (node-elem bt)
               (append (tree-elems (node-left bt))
                       (tree-elems (node-right bt))))]))
\end{verbatim}
Here is an abstraction function for trees that is similar to
{\tt foldr} for lists, but works on trees:
\begin{verbatim}
;; tree-fold : [X Y] . [X Y Y -> Y] Y [Tree X] -> Y
;; The fundamental abstraction function for trees
(define (tree-fold f b bt)
  (cond [(leaf? bt) b]
        [(node? bt)
         (f (node-elem bt)
            (tree-fold f b (node-left bt))
            (tree-fold f b (node-right bt)))]))
\end{verbatim}
Give an equivalent definition of {\tt tree-elems} in terms of {\tt
  tree-fold}.  (Just provide the code.)

(define (tree-elems bt)
  (tree-fold (lambda (x l r) (cons x (append l r))) '() bt))

\newpage
\noindent
{\bf Problem 5 (15 points).}  
%
Design a program that takes a list of strings and produces the count
of strings with length longer than 3 in the list.  For example, if the
list contains {\tt "a"}, {\tt "dave"}, {\tt "abc"}, and {\tt "fred"},
the count is 2.  You may assume the {\tt{[Listof String]}} data
definition is defined.
%
For full credit, use list abstraction functions.  For partial credit,
follow the template for {\tt{[Listof String]}}.

;; count-3+ : [Listof String] -> Number
;; Count the number of strings with length more than 3 in given list
(check-expect (count-3+ '()) 0)
(check-expect (count-3+ (list "a" "abcd" "fred" "c")) 2)
(define (count-3+ los)
  (foldr (lambda (s c) (if (> (string-length s) 3) (add1 c) c)) 0 los))

;; Alts:
(define (count-3+ los)
  (local [;; count+ : String Number -> Number
          ;; Bump count if string has length > 3
          (define (count+ s c) 
            (if (> (string-length s) 3) (add1 c) c))]
    (foldr count+ 0 los))

(define (count-3+ los)
  (local [;; 3+? : String -> Boolean
          ;; Does the string have length > 3?
          (define (3+? s) 
            (> (string-length s) 3))]
    (length (filter 3+? los))))

(define (count-3+ los)
  (length (filter (lambda (s) (> (string-length s) 3)) los)))
\end{SaveVerbatim}

@@

(define (map f xs)
  (foldr (lambda (x ys) (cons (f x) ys)) '() xs))
\end{SaveVerbatim}

@@

\noindent
{\bf Problem 7 (15 points).}
%
Here's a data definition for choose-your-own-adventure books:
\begin{verbatim}
;; A CYOA is one:
;; - String
;; - (make-choice String CYOA CYOA)
(define-struct choice (q yes no))
;; Interp: a choose-your-own-adventure book where a string
;; is a conclusion and a choice is a yes/no question and
;; the two adventures that follow as a consequnce of the answer.
\end{verbatim}

\noindent
Design a function called {\tt ans} that consumes a {\tt CYOA} and a
{\tt [Listof Boolean]} that represents answers to the yes/no
questions: {\tt \#true} means yes, {\tt \#false} means no.  The {\tt
  ans} function produces the {\tt CYOA} adventure that results after
giving all of the answers in the list.  If the list contains too many
answers, {\tt ans} just produces the conclusion of the adventure.

;; most-games : Tournament -> Number
;; Compute most games any team has to win to win tournament
(check-expect (most-games "NY") 0)
(check-expect (most-games (make-bracket "LA" "HOU")) 1)
(check-expect (most-games (make-bracket (make-bracket "LA" "DC") "HOU")) 2)
(define (most-games t)
  (cond [(string? t) 0]
        [(bracket? t)
         (add1 (max (most-games (bracket-t1 t))
                    (most-games (bracket-t2 t))))]))
\end{SaveVerbatim}
